---
title: "e1 adult analysis"
output: pdf_document
date: "2025-03-12"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
library(tidyverse)
library(here)
library(lme4)
library(broom.mixed)
library(tidyboot)
library(svglite)
library(ggthemes)
library(ggrepel)
```

```{r, eval = FALSE}
raw_data <- list.files( path=here("../data/e1_adult_raw_data"), full.names=TRUE ) %>% 
  map_dfr( read_csv )

data <- raw_data %>%
  filter("study_id" != "UNSAVED_STUDY",
         !is.na(study_id)) %>%
  select(-prolific_id) 

write_csv(data, here("../data/e1_adult_data.csv"))
```

```{r}
data <- read_csv(here("../data/e1_adult_data.csv")) %>%
  mutate(stimulus_subclass = str_remove(image_name, "^[^_]*_"),
         stimulus_subclass = str_remove(stimulus_subclass, "_.*")) %>%
  mutate(typicality = typicality - 1)

p <- data %>%
  group_by(condition, participant_id) %>%
  mutate(prop_atypical = sum(typicality)/n()) %>%
  ungroup() %>%
  group_by(condition) %>%
  tidyboot_mean(prop_atypical) %>%
  ungroup() %>%
  ggplot(aes(x = condition, y = mean)) + 
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper)) +
  ylab("proportion atypical choices") +
  theme_few() + theme(
  panel.background = element_rect(fill = "transparent",
                                  colour = NA_character_), # necessary to avoid drawing panel outline
  panel.grid.major = element_blank(), # get rid of major grid
  panel.grid.minor = element_blank(), # get rid of minor grid
  plot.background = element_rect(fill = "transparent",
                                 colour = NA_character_), # necessary to avoid drawing plot outline
  legend.background = element_rect(fill = "transparent"),
  legend.box.background = element_rect(fill = "transparent"),
  legend.key = element_rect(fill = "transparent")
)

ggsave(
  plot = p,
  filename = here("plots/pilot_typicality_effect.svg"),
  height = 5,
  width = 5
)
```

Do people tend to choose more atypical objects when prompted with a novel word?

Participants' choices on familiar word (e.g., "leaf") vs. novel word (e.g., "neft") trials. Higher values indicate they chose more atypical items. Stimuli typicality scored by experimenter on a binary scale. People choose more atypical exemplars when prompted with a novel compared to a familiar word. They also strongly prefer typical exemplars when prompted with familiar words.

```{r}
glmer(typicality ~ condition + (1|participant_id) + (1 | category), family = "binomial",
      data = data) %>%
  tidy()
```

```{r}
data <- data %>%
  group_by(participant_id, trial_number) %>%
  mutate(distinct_types = n_distinct(stimulus_subclass) - 1) %>%
  ungroup() 

p <- data %>%
  group_by(condition, participant_id, trial_number) %>%
  mutate(prop_distinct = sum(distinct_types)/n()) %>%
  ungroup() %>%
  group_by(condition) %>%
  tidyboot_mean(prop_distinct) %>%
  ungroup() %>%
  ggplot(aes(x = condition, y = mean)) + 
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper)) +
  ylab("proportion choices of multiple types") +
  theme_few() + theme(
  panel.background = element_rect(fill = "transparent",
                                  colour = NA_character_), # necessary to avoid drawing panel outline
  panel.grid.major = element_blank(), # get rid of major grid
  panel.grid.minor = element_blank(), # get rid of minor grid
  plot.background = element_rect(fill = "transparent",
                                 colour = NA_character_), # necessary to avoid drawing plot outline
  legend.background = element_rect(fill = "transparent"),
  legend.box.background = element_rect(fill = "transparent"),
  legend.key = element_rect(fill = "transparent")
)

ggsave(
  plot = p,
  filename = here("plots/pilot_types_effect.svg"),
  height = 5,
  width = 5
)
  
```

Do people tend to choose more varied objects when prompted with familiar vs. novel words? Another way to tell whether people think the meaning of a novel word is narrower is to test whether they choose one subtype (e.g., bean-shaped leaves) vs. multiple subtypes (e.g., teardrop-shaped and oak-shaped leaves). 

Above, participants' choices of exemplars; they have two choices on each trial. A score of 0 reflects that they chose all of one subtype, and a score of 1 reflects that they chose two distinct subtypes. People tend to choose fewer distinct subtypes when prompted with novel words.

```{r}
glmer(distinct_types ~ condition + (1|participant_id) + (1 | category), family = "binomial",
      data = data) %>%
  tidy()
```

```{r}
p <- data %>%
  group_by(condition, participant_id, category) %>%
  mutate(prop_atypical = sum(typicality)/n()) %>%
  ungroup() %>%
  group_by(condition, category) %>%
  tidyboot_mean(prop_atypical) %>%
  ungroup() %>%
  ggplot(aes(x = condition, y = mean)) + 
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper)) +
  facet_wrap(~category) +
  ylab("proportion atypical choices") +
  theme_few() + theme(
  panel.background = element_rect(fill = "transparent",
                                  colour = NA_character_), # necessary to avoid drawing panel outline
  panel.grid.major = element_blank(), # get rid of major grid
  panel.grid.minor = element_blank(), # get rid of minor grid
  plot.background = element_rect(fill = "transparent",
                                 colour = NA_character_), # necessary to avoid drawing plot outline
  legend.background = element_rect(fill = "transparent"),
  legend.box.background = element_rect(fill = "transparent"),
  strip.background =element_rect(fill="transparent"),
  legend.key = element_rect(fill = "transparent")
)

ggsave(
  plot = p,
  filename = here("plots/pilot_atypical_by_item_effect.svg"),
  bg = "transparent",
  height = 5,
  width = 7
)
```

Examining typicality across stimulus types.

```{r}
p <- data %>%
  group_by(condition, participant_id, category, trial_number) %>%
  mutate(prop_distinct = sum(distinct_types)/n()) %>%
  ungroup() %>%
  group_by(condition, category) %>%
  tidyboot_mean(prop_distinct) %>%
  ungroup() %>%
  ggplot(aes(x = condition, y = mean)) + 
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper)) +
  facet_wrap(~category) +
  ylab("proportion choices of multiple distinct subtypes") +
  theme_few() + theme(
  panel.background = element_rect(fill = "transparent",
                                  colour = NA_character_), # necessary to avoid drawing panel outline
  panel.grid.major = element_blank(), # get rid of major grid
  panel.grid.minor = element_blank(), # get rid of minor grid
  plot.background = element_rect(fill = "transparent",
                                 colour = NA_character_), # necessary to avoid drawing plot outline
  legend.background = element_rect(fill = "transparent"),
  legend.box.background = element_rect(fill = "transparent"),
  strip.background =element_rect(fill="transparent"),
  legend.key = element_rect(fill = "transparent")
)

ggsave(
  plot = p,
  filename = here("plots/pilot_types_by_item_effect.png"),
  bg = "transparent",
  height = 5,
  width = 7
)
```

Examining distinct type choices across stimulus types.

```{r, eval = FALSE}

adult_data <- adult_data %>%
  mutate(age = "Adults") %>%
  select(-study_id, -session_id)

plot_data <- data %>%
  filter(category != "carrots", category != "apples") %>%
  group_by(participant_id, trial_number) %>%
  mutate(distinct_types = n_distinct(stimulus_subclass) - 1) %>%
  ungroup() %>%
  select(participant_id, trial_number, condition, category, image_name, word, click_order, rt, id, typicality,
         stimulus_subclass, distinct_types) %>%
  mutate(age = "3- to 5-year-olds") %>%
  rbind(adult_data) 

p <- plot_data %>%
  group_by(age, condition, participant_id, trial_number) %>%
  summarise(prop_distinct = sum(distinct_types)/n()) %>%
  ungroup() %>%
  group_by(age, condition) %>%
  tidyboot_mean(prop_distinct) %>%
  ungroup() %>%
  ggplot(aes(x = condition, y = mean, color = age)) + 
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper),
                  position = position_dodge(width = 0.2)) +
  geom_line(aes(group = age),
            position = position_dodge(width = 0.2)) +
  ylab("proportion choices of multiple types") +
  scale_color_brewer(palette = "Dark2") +
  theme_few() + theme(
  panel.background = element_rect(fill = "transparent",
                                  colour = NA_character_), # necessary to avoid drawing panel outline
  panel.grid.major = element_blank(), # get rid of major grid
  panel.grid.minor = element_blank(), # get rid of minor grid
  plot.background = element_rect(fill = "transparent",
                                 colour = NA_character_), # necessary to avoid drawing plot outline
  legend.background = element_rect(fill = "transparent"),
  legend.box.background = element_rect(fill = "transparent"),
  legend.key = element_rect(fill = "transparent")
) +
  geom_text_repel(
    data = . %>%
      group_by(age) %>%
      filter(condition == max(condition)) %>%  # assumes 'condition' is numeric or ordered factor
      ungroup(),
    aes(label = age),
    nudge_x = 0.1,
    show.legend = FALSE,
    direction = "y",
    hjust = 0,
    segment.color = NA
  )

ggsave(
  plot = p,
  filename = here("plots/overall_types_effect.svg"),
  height = 5,
  width = 6
)

p <- plot_data %>%
  group_by(age, condition, participant_id) %>%
  summarise(prop_atypical = sum(typicality)/n()) %>%
  ungroup() %>%
  group_by(age, condition) %>%
  tidyboot_mean(prop_atypical) %>%
  ungroup() %>%
  ggplot(aes(x = condition, y = mean, color = age)) + 
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper),
                  position = position_dodge(width = 0.2)) +
  geom_line(aes(group = age), position = position_dodge(width = 0.2)) +
  ylab("proportion atypical choices") +
  scale_color_brewer(palette = "Dark2") +
  theme_few() + theme(
  panel.background = element_rect(fill = "transparent",
                                  colour = NA_character_), # necessary to avoid drawing panel outline
  panel.grid.major = element_blank(), # get rid of major grid
  panel.grid.minor = element_blank(), # get rid of minor grid
  plot.background = element_rect(fill = "transparent",
                                 colour = NA_character_), # necessary to avoid drawing plot outline
  legend.background = element_rect(fill = "transparent"),
  legend.box.background = element_rect(fill = "transparent"),
  legend.key = element_rect(fill = "transparent")
) +
  geom_text_repel(
    data = . %>%
      group_by(age) %>%
      filter(condition == max(condition)) %>%  # assumes 'condition' is numeric or ordered factor
      ungroup(),
    aes(label = age),
    nudge_x = 0.1,
    show.legend = FALSE,
    direction = "y",
    hjust = 0,
    segment.color = NA
  )

ggsave(
  plot = p,
  filename = here("plots/overall_typicality_effect.svg"),
  height = 5,
  width = 6
)

```